---
title: "R Notebook for Emotion Rating Database Analysis"
author: "Ari Dyckovsky"
output:
  pdf_document: default
  html_notebook: default
---

# Emotionn Rating Analysis

## Load packages

```{r}
library(dplyr)
library(purrr)
library(tidyr)
library(stringr)
library(ggplot2)
library(hash)
library(scales)
library(readr)
```

## Constants

```{r}
# Prevalence measures
HOMOSEXUAL_PREVALENCE <- 0.1
HETEORSEXUAL_PREVALENCE <- 0.9
HOMOSEXUAL_TAG <- "_ho"

# Database details
DB_SCALES <- hash()
DB_SCALES[["OASIS"]] <- c(1, 7)
DB_SCALES[["IAPS"]] <- c(1, 9)
DB_SCALES[["NAPS_BE"]] <- c(1, 9)
DB_SCALES[["NAPS_ERO"]] <- c(1, 9)

# Output rating scale
POS_NEG_RANGE <- c(-1, 1)

# For formatting doubles
options(digits = 2)
```

## Set datapath and load `shlab.imgct`

Currently pulling from local, but will make dynamic for other machines later.

```{r}
# Set the working directory to be part of S Drive (may make dynamic later?)
# Whilst not dynamic, change for own session if mount point is not equivalent on
# local machine
shared_dir <- "~/Projects/shlab/mounts/imgct"
package_dir <- "~/Projects/shlab"

datapath <- file.path(shared_dir, "csn_images")
imgct_package_path <- file.path(package_dir, "shlab.imgct")

# Make sure that devtools, tidyverse are installed before this call
devtools::load_all(imgct_package_path)
```

## Load Database Imports

```{r}
oasis_df <- shlab.imgct::load_imported_xlsx(datapath, "oasis_emotion_ratings")
iaps_df <- shlab.imgct::load_imported_xlsx(datapath, "iaps_emotion_ratings")
naps_be_df <- shlab.imgct::load_imported_xlsx(datapath, "naps_be_emotion_ratings")
naps_ero_df <- shlab.imgct::load_imported_xlsx(datapath, "naps_ero_emotion_ratings")
```

### Handle Lack of General Ratings for NAPS ERO dataset

Provided that NAPS ERO has specified arousal and valence ratings for four groups, but no general rating of arousal and valence, we must choose an averaging strategy. We will attempt a weighted average with rough prevelance estimates, based on these four groups:

1. HoF (Homosexual Female)
2. HoM (Homosexual Male)
3. HeF (Heterosexual Female)
4. HeM (Heterosexual Male)

Constants for prevalence are determined above, and below is a function for weighted ratings based upon the columns of NAPS ERO where "_ho" and "_he" are tags for homosexual and heterosexual, respectively.

```{r}
weighted_rating <- function(ratings) {
  columns <- names(ratings)
  
  # Weights are distributed to tags such that:
  #   - FALSE = HETEROSEXUAL_PREVALENCE
  #   - TRUE = HOMOSEXUAL_PREVALENCE
  weights <- c(
    HETEORSEXUAL_PREVALENCE,
    HOMOSEXUAL_PREVALENCE
  )
  
  tags <- stringr::str_detect(columns, HOMOSEXUAL_TAG)
  weights_vec <- weights[1 + tags]

  weighted.mean(
    ratings,
    weights_vec
  )
}
```

Mutate new columns of general arousal and valence ratings for NAPS ERO.

```{r}
naps_ero_df <- naps_ero_df %>%
  dplyr::mutate(
    arousal = dplyr::select(., starts_with("arousal")) %>%
      purrr::pmap_dbl(
        .,
        ~ weighted_rating(c(...))
      ),
    valence = dplyr::select(., starts_with("valence")) %>%
      purrr::pmap_dbl(
        .,
        ~ weighted_rating(c(...))
      )
  ) %>%
  dplyr::relocate(
    c(arousal, valence),
    .after = image_id
  )

naps_ero_df
```

```{r}
naps_ero_df %>% summarize(mean = mean(arousal))
```


## Merge dataframes

```{r}
rescale_rating <- function(rating, db) {
  db_scale <- DB_SCALES[[db]]
  return(scales::rescale(rating, from = db_scale, to = POS_NEG_RANGE))
}

all_emotion_ratings_df <- dplyr::bind_rows(
    list(
      oasis_df,
      iaps_df,
      naps_be_df,
      naps_ero_df
    ),
    .id = "db_id"
  ) %>%
  dplyr::select(db_id, image_id, arousal, valence) %>%
  dplyr::mutate_at(
    vars(db_id),
    ~ dplyr::recode(
      .,
      "1" = "OASIS",
      "2" = "IAPS",
      "3" = "NAPS_BE",
      "4" = "NAPS_ERO"
    )
  ) %>%
  dplyr::mutate(
    arousal = pmap_dbl(list(arousal, db_id), rescale_rating),
    valence = pmap_dbl(list(valence, db_id), rescale_rating)
  )

all_emotion_ratings_df
```

```{r}
p <- ggplot(all_emotion_ratings_df, aes(x=valence, y=arousal, color=db_id)) +
  geom_point(alpha = 0.5) +
  scale_x_continuous(limits = POS_NEG_RANGE) +
  scale_y_continuous(limits = POS_NEG_RANGE)

p <- p + labs(
    title = "Arousal vs. Valence of Database Images",
    x = "Valence",
    y = "Arousal",
    color = "Source"
  ) +
  theme_minimal() +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0)

p
```


## Save to results as general TSV of arousal and valence ratings

```{r}
readr::write_tsv(
  all_emotion_ratings_df,
  file.path(datapath, "results", "general_emotion_ratings.tsv"),
  append = FALSE,
  col_names = TRUE
)
```

## Random checks on data

```{r}
all_emotion_ratings_df %>%
  dplyr::group_by(db_id) %>%
  dplyr::count()
```

