---
title: "Validate Response Data"
author: "Ari Dyckovsky"
output:
  md_document:
    variant: markdown_github
---

# Validate Response Data

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  strip.white = TRUE,
  tidy = TRUE,
  highlight = TRUE
)
```

## Load packages

```{r, package-loading, message=FALSE}
library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
library(hash)
```

## Set datapath and load `shlab.imgct`

Begin by setting the working directory and important top-level paths to data and 
loading necessary packages.

- NOTE: This will be changed to dynamically account for the package `shlab.imgct` via
  its GitHub instance later. For now, it is using development loading.

```{r, shlab-setup, message=FALSE}
# Set the working directory to be part of S Drive (may make dynamic later?)
# Whilst not dynamic, change for own session if mount point is not equivalent on
# local machine
shared_dir <- "~/Projects/shlab/mounts/imgct"
package_dir <- "~/Projects/shlab"

datapath <- file.path(shared_dir, "csn_images")
imgct_package_path <- file.path(package_dir, "shlab.imgct")

# Make sure that devtools, tidyverse are installed before this call
devtools::load_all(imgct_package_path)
```

## Validation of All Clean Blocks

The `shlab.imgct` package provides a comprehensive validation method to load and validate all clean blocks of participant responses against a "validation key". This comprehensive method is called `validate_all_participants`. It is a convenient abstraction on top of `load_all_clean_blocks`, `load_key`, `validate_block_participants`, which can be used individually. Below, we first show the use of these underlying methods, but will use the abstraction to handle the data for later analysis needs.

### Load All Clean Blocks

```{r, load-clean-blocks}
all_clean_blocks <- shlab.imgct::load_all_clean_blocks(datapath)
```

Example of a loaded clean block by selected the first of the list of loaded dataframes.

```{r, example-clean-block}
one_clean_block <- all_clean_blocks[[1]]

knitr::kable(
  one_clean_block
)
```

### Load the Validation Key

This method loads a TXT key that is used to compare participant responses to images specified for validation during the task.

```{r, load-validation-key}
validation_key <- shlab.imgct::load_key(datapath, "validation_key.txt")
```

What does this key look like when loaded?

```{r, example-validation-key}
knitr::kable(
  validation_key
)
```

### Validate a Clean Block

Here, we show how to validate a single clean block of participant responses against the validation key.

```{r, validate-one-clean-block}
one_validated_block <- shlab.imgct::validate_block_participants(one_clean_block, validation_key)

knitr::kable(
  one_validated_block
)
```

### Validate All Participants

Using a similar procedure as above, all clean blocks of participant responses can be validated by `validate_all_participants`, which will also save a TSV to file.

```{r, results='hide'}
shlab.imgct::validate_all_participants(datapath, key_filename = "validation_key.txt")
```
